Values disclosures and trust in science: A replication study

### 2022-10-18 paper/paper.Qmd
2 grammar issues

As its name indicates, METI assesses perceived trustworthiness rather than trust (such as accepting Dr. Spence's conclusion about BPA).  It is conceivable that participants might trust Dr. Spence's conclusion even if they judge Spence to be untrustworthy, or vice versa.  But we would typically trust and trustworthiness to go together.

@@

All together, regarding the philosophical debates over values in science, the evidence from this replication study suggests that the effects of transparency on perceived trustworthiness might be context-dependent.  Transparency about the influence of might be detrimental in some cases, beneficial in others, and neutral in still others.  Notably, both of the papers cited as objecting to transparency [@JohnEpistemicTrustEthics2017; @KovakaClimateChangeDenial2021] focus on climate change, while many of Elliott's examples come from toxicology or environmental public health.  The critics might be right that transparency has undermined trust in the climate context, while Elliott might be right that transparency can be neutral or even beneficial in some public health contexts.  Because philosophers of science tend to work with a small number of rich case studies, methods such as qualitative comparative analysis [QCA\; @RaginComparativeMethodMoving2013] would be worth exploring.

@@

In more prominent and/or polarized cases than BPA, a number of other psychological mechanisms might be activated.  Motivated reasoning or identity-protective cognition [@KahanPolarizingImpactScience2012] might cause members of the general public to rate a scientist as more or less trustworthy insofar as the scientist agrees with the respondent's "side" of the debate, regardless of whether and which values are disclosed.  That is, the scientist conclusion might have a larger effect than scientist values.  Alternatively, respondents might use values disclosures to identify which scientists are on "their side," and so scientist values would have a larger effect than scientist conclusion.  Careful attention to experimental design will be needed to separate these potential multiway interactions between conclusions, disclosures, and public prominence.

>>>>

As its name indicates, METI assesses perceived trustworthiness rather than trust (such as accepting Dr. Spence's conclusion about BPA).  It is conceivable that participants might trust Dr. Spence's conclusion even if they judge Spence to be untrustworthy, or vice versa.  But we would typically expect trust and trustworthiness to go together.

@@

All together, regarding the philosophical debates over values in science, the evidence from this replication study suggests that the effects of transparency on perceived trustworthiness might be context-dependent.  Transparency about the influence of values might be detrimental in some cases, beneficial in others, and neutral in still others.  Notably, both of the papers cited as objecting to transparency [@JohnEpistemicTrustEthics2017; @KovakaClimateChangeDenial2021] focus on climate change, while many of Elliott's examples come from toxicology or environmental public health.  The critics might be right that transparency has undermined trust in the climate context, while Elliott might be right that transparency can be neutral or even beneficial in some public health contexts.  Because philosophers of science tend to work with a small number of rich case studies, methods such as qualitative comparative analysis [QCA\; @RaginComparativeMethodMoving2013] would be worth exploring.

@@

In more prominent and/or polarized cases than BPA, a number of other psychological mechanisms might be activated.  Motivated reasoning or identity-protective cognition [@KahanPolarizingImpactScience2012] might cause members of the general public to rate a scientist as more or less trustworthy insofar as the scientist agrees with the respondent's "side" of the debate, regardless of whether and which values are disclosed.  That is, the scientist's conclusion might have a larger effect than scientist's disclosed values.  Alternatively, respondents might use values disclosures to identify whether a scientist is on "their side," and so the scientist's disclosed values might have a larger effect than the scientist's conclusion.  Careful attention to experimental design will be needed to separate these potential multiway interactions between conclusions, disclosures, and public prominence.

### 2022-10-18 paper/paper.tex
2 grammar issues

\begin{document}
\maketitle
\begin{abstract}
While philosophers of science generally agree that social, political,
and ethical values can play legitimate roles in science, there is active
debate over whether scientists should disclosure such values in their
public communications. This debate depends, in part, on empirical claims
about whether values disclosures might undermine public trust in
science. In a previous study, Elliott et al.~used an online experiment
to test this empirical claim. The current paper reports a replication
attempt of their experiment. Comparing results of the original study and
our replication, we do not find evidence for a transparency penalty or
``shared values'' effect, but do find evidence that the content of
scientific conclusions (whether or not a chemical is found to cause
harm) might effect perceived trustworthiness and that scientists who
value public health and disclose this value might be perceived as more
trustworthy.
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[sharp corners, enhanced, frame hidden, breakable, borderline west={3pt}{0pt}{shadecolor}, boxrule=0pt, interior hidden]}{\end{tcolorbox}}\fi

@@

As its name indicates, METI assesses perceived trustworthiness rather
than trust (such as accepting Dr.~Spence's conclusion about BPA). It is
conceivable that participants might trust Dr.~Spence's conclusion even
if they judge Spence to be untrustworthy, or vice versa. But we would
typically trust and trustworthiness to go together.

@@

All together, regarding the philosophical debates over values in
science, the evidence from this replication study suggests that the
effects of transparency on perceived trustworthiness might be
context-dependent. Transparency about the influence of might be
detrimental in some cases, beneficial in others, and neutral in still
others. Notably, both of the papers cited as objecting to transparency
(John 2017; Kovaka 2021) focus on climate change, while many of
Elliott's examples come from toxicology or environmental public health.
The critics might be right that transparency has undermined trust in the
climate context, while Elliott might be right that transparency can be
neutral or even beneficial in some public health contexts. Because
philosophers of science tend to work with a small number of rich case
studies, methods such as qualitative comparative analysis (QCA; Ragin
2013) would be worth exploring.

@@

In more prominent and/or polarized cases than BPA, a number of other
psychological mechanisms might be activated. Motivated reasoning or
identity-protective cognition (Kahan et al. 2012) might cause members of
the general public to rate a scientist as more or less trustworthy
insofar as the scientist agrees with the respondent's ``side'' of the
debate, regardless of whether and which values are disclosed. That is,
the scientist conclusion might have a larger effect than scientist
values. Alternatively, respondents might use values disclosures to
identify which scientists are on ``their side,'' and so scientist values
would have a larger effect than scientist conclusion. Careful attention
to experimental design will be needed to separate these potential
multiway interactions between conclusions, disclosures, and public
prominence.

>>>>

\begin{document}
\maketitle
\begin{abstract}
While philosophers of science generally agree that social, political,
and ethical values can play legitimate roles in science, there is active
debate over whether scientists should disclosure such values in their
public communications. This debate depends, in part, on empirical claims
about whether values disclosures might undermine public trust in
science. In a previous study, Elliott et al.~used an online experiment
to test this empirical claim. The current paper reports a replication
attempt of their experiment. Comparing results of the original study and
our replication, we do not find evidence for a transparency penalty or
``shared values'' effect, but do find evidence that the content of
scientific conclusions (whether or not a chemical is found to cause
harm) might effect perceived trustworthiness and that scientists who
value public health and disclose this value might be perceived as more
trustworthy.
\end{abstract}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, interior hidden, enhanced, boxrule=0pt, sharp corners, breakable, frame hidden]}{\end{tcolorbox}}\fi

@@

As its name indicates, METI assesses perceived trustworthiness rather
than trust (such as accepting Dr.~Spence's conclusion about BPA). It is
conceivable that participants might trust Dr.~Spence's conclusion even
if they judge Spence to be untrustworthy, or vice versa. But we would
typically expect trust and trustworthiness to go together.

@@

All together, regarding the philosophical debates over values in
science, the evidence from this replication study suggests that the
effects of transparency on perceived trustworthiness might be
context-dependent. Transparency about the influence of values might be
detrimental in some cases, beneficial in others, and neutral in still
others. Notably, both of the papers cited as objecting to transparency
(John 2017; Kovaka 2021) focus on climate change, while many of
Elliott's examples come from toxicology or environmental public health.
The critics might be right that transparency has undermined trust in the
climate context, while Elliott might be right that transparency can be
neutral or even beneficial in some public health contexts. Because
philosophers of science tend to work with a small number of rich case
studies, methods such as qualitative comparative analysis (QCA; Ragin
2013) would be worth exploring.

@@

In more prominent and/or polarized cases than BPA, a number of other
psychological mechanisms might be activated. Motivated reasoning or
identity-protective cognition (Kahan et al. 2012) might cause members of
the general public to rate a scientist as more or less trustworthy
insofar as the scientist agrees with the respondent's ``side'' of the
debate, regardless of whether and which values are disclosed. That is,
the scientist's conclusion might have a larger effect than scientist's
disclosed values. Alternatively, respondents might use values
disclosures to identify whether a scientist is on ``their side,'' and so
the scientist's disclosed values might have a larger effect than the
scientist's conclusion. Careful attention to experimental design will be
needed to separate these potential multiway interactions between
conclusions, disclosures, and public prominence.
